---
title: "House Prices in Portland, OR"
author: "Karol Orozco"
date: "2022-12"
---

## Background

The goal is to build a classification model to predict the type of median housing prices in Portland, OR and its metropolitan area.

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(tidymodels)
library(dplyr)
library(gt) ## tables
library(ggmap)



suppressWarnings(if(!require("pacman")) install.packages("pacman"))

pacman::p_load('tidyverse', 'tidymodels', 'ranger','vip', 'ggplot2', 'readr', 'ggmap','gt' )
```

## Get the Data

```{r}
raw_pdx <- read.csv("https://raw.githubusercontent.com/karolo89/Raw_Data/main/PORTLAND%20HOUSE.csv", stringsAsFactors=TRUE)
```

## Prepare the Data

This data has 25731 obs. of 32 variables

```{r}
## raw_pdx <- raw_pdx%>%select(-id)

head(raw_pdx)
```

```{r}

# convert variables

raw_pdx <-  raw_pdx %>% 
  
  mutate(
    
    yearBuilt = as.numeric(yearBuilt),
    bathrooms = as.numeric(bathrooms),
    bedrooms = as.numeric(bedrooms),
    daysOnZillow = as.numeric(daysOnZillow),
    lastSoldPrice = as.numeric(lastSoldPrice ),
    livingArea = as.numeric(livingArea),
    lotSize= as.numeric(lotSize),
    price = as.numeric(price),
    priceHistory.1.price= as.numeric(priceHistory.1.price),
    
    
    ElementarySchoolrating = as.factor(ElementarySchoolrating),
    MiddleSchoolsrating = as.factor(MiddleSchoolsrating),
    HighSchoolRating= as.factor(HighSchoolRating),
    zipcode = as.factor(zipcode)

    
  )

```

### Missing data

```{r}
is.na(raw_pdx) %>% colSums()
```

```{r}
clean_data <- raw_pdx %>%
  filter(!is.na(yearBuilt))%>%
  filter(!is.na(longitude))%>%
  filter(!is.na(bedrooms))%>%
  filter(!is.na(daysOnZillow))%>%
  filter(!is.na(livingArea))%>%
  filter(!is.na(priceHistory.1.price))%>%
  filter(!is.na(hasFireplace))%>%
  filter(!is.na(latitude))%>%
  filter(!is.na(hasHeating))%>%
  filter(!is.na(hasCooling))%>%
  filter(!is.na(bathrooms))%>%
  filter(!is.na(lotSize))%>%
  filter(!is.na(propertyTaxRate))%>%
  filter(!is.na(ElementarySchooldistance))%>%
  filter(!is.na(MiddleSchooldistance))%>%
  filter(!is.na(HighSchooldistance))%>%
  filter(!is.na(ElementarySchoolrating))%>%
  filter(!is.na(MiddleSchoolsrating))%>%
  filter(!is.na(HighSchoolRating))

```

```{r}
summary(clean_data)

```

```{r}
clean_data <- 
  clean_data %>% 
  mutate(price_category = case_when( 
    price < 551000 ~ "below",
    price >= 551000 ~ "above")) %>% 
  mutate(price_category = as.factor(price_category)) 
```

## Take a look at the Data

```{r}
clean_data %>% 
  count(price_category, 
        name ="total") %>%
  mutate(percent = total/sum(total)*100,
         percent = round(percent, 2)) %>%
 gt() %>%
  tab_header(
    title = "Portland, OR and its Metropolitan Area Median House Prices",
    subtitle = "Above and below 551,000$"
  ) %>%
  cols_label(
    price_category = "Price",
    total = "Total",
    percent = "Percent"
  ) %>% 
  fmt_number(
    columns = vars(total),
    suffixing = TRUE
  )  
```

```{r}
qmplot(x = longitude, 
       y = latitude, 
       data = clean_data, 
       geom = "point", 
       color = price_category, 
       alpha = 0.4) +
  scale_alpha(guide = 'none')

```

```{r}
houses_pdx <-
  clean_data %>% 
  select( # select our predictors
    longitude, 
    latitude, 
    price_category,
    bathrooms, 
    yearBuilt, 
    homeType,
    bedrooms, 
    livingArea, 
    lotSize,
    MiddleSchooldistance,
    ElementarySchooldistance,
    HighSchooldistance)

glimpse(houses_pdx)

```

```{r}
pdx_long <- houses_pdx %>% 
  select(-longitude,-latitude, -homeType, -yearBuilt, -lotSize)%>%
    pivot_longer(!price_category, names_to = "features", values_to = "values")


# Print the first 10 rows
pdx_long %>% 
  slice_head(n = 10)
```

```{r}
theme_set(theme_light())

# Make a box plot for each predictor feature
pdx_long %>% 
  ggplot(mapping = aes(x = price_category, y = values, fill = features)) +
  geom_boxplot() + 
  facet_wrap(~ features, scales = "free", ncol = 4) +
  scale_color_viridis_d(option = "plasma", end = .7) +
  theme(legend.position = "none")
```

### Data Splitting

```{r}

# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible 
set.seed(504)

# Put 3/4 of the data into the training set 
data_split <- initial_split(houses_pdx, 
                           prop = 3/4)

# Create dataframes for the two sets:
train_data <- training(data_split) 
test_data <- testing(data_split)
```

### Validaton Set

```{r}
house_folds <-
 vfold_cv(train_data, 
          v = 5, 
          strata = price_category) 
```

```{r}
pdx_rec <-
  recipe(price_category ~ .,
         data = train_data) %>%
  update_role(longitude, latitude, 
              new_role = "ID") %>% 
  
  step_naomit(everything(), skip = TRUE) %>% 
  
  step_novel(all_nominal(), -all_outcomes()) %>% # converts all nominal variables to factors and takes care of other issues related to categorical variables.
  
  step_normalize(all_numeric(), -all_outcomes(), 
                 -longitude, -latitude) %>% # step_normalize() normalizes (center and scales) the numeric variables to have a standard deviation of one and a mean of zero
  
  step_dummy(all_nominal(), -all_outcomes()) %>% #converts our factor columns into numeric binary (0 and 1) variables.
  
  step_zv(all_numeric(), -all_outcomes()) %>% ## step_zv(): removes any numeric variables that have zero variance.
  
  step_corr(all_predictors(), threshold = 0.7, method = "spearman") # step_corr(): will remove predictor variables that have large correlations with other predictor variables.

```

```{r}
summary(pdx_rec)

```

```{r}
prep_data <- 
  pdx_rec %>% # use the recipe object
  prep() %>% # perform the recipe on training data
  juice() # extract only the preprocessed dataframe 

```

## The Model- Logistic regression

```{r}
log_spec <- # your model specification
  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec
```

```{r}

pdx_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(pdx_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

pdx_wflow

```

```{r, message=FALSE, warning=FALSE}
# save model coefficients for a fitted model object from a workflow

get_model <- function(x) {
  pull_workflow_fit(x) %>% tidy()
}

# same as before with one exception
log_res_2 <- 
  pdx_wflow %>% 
  fit_resamples(
    resamples = house_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
      control = control_resamples(
      save_pred = TRUE,
      extract = get_model) # use extract and our new function
    ) 


## All of the results can be flattened and collected using:



all_coef <- map_dfr(log_res_2$.extracts, ~ .x[[1]][[1]])
filter(all_coef, term == "bedrooms")

```

### Performance metrics

Show performance for every single fold:

```{r}
log_res_2 %>%  collect_metrics(summarize = FALSE)

```

### Collect predictions

To obtain the actual model predictions, we use the function collect_predictions and save the result as log_pred:

```{r}
log_pred <- 
  log_res_2 %>%
  collect_predictions()


log_pred %>% 
  conf_mat(price_category, .pred_class) 
```

```{r}
log_pred %>% 
  conf_mat(price_category, .pred_class) %>% 
  autoplot(type = "heatmap")+
  theme_minimal()
```

### ROC Curve

```{r}
log_pred %>% 
  group_by(id) %>% # id contains our folds
  roc_curve(price_category, .pred_above) %>% 
  autoplot()+
  theme_minimal()
```

## Use the workflow to train our model

```{r}
pdx_fit <- fit(pdx_wflow, train_data)
```

This allows us to use the model trained by this workflow to predict labels for our test set, and compare the performance metrics with the basic model we created previously.

```{r}
pdx_fit %>% ## display results
pull_workflow_fit() %>%
tidy()%>%
  filter(p.value < 0.05)
```

```{r}
library(vip)

pdx_fit %>%
extract_fit_parsnip() %>%
   vip(num_features = 5)+
  theme_minimal()
```

The two most important predictors in whether the median house value is above or below 551,000 dollars were the Living Area and the home type: Townhouse

```{r}
# Make predictions on the test set
pred_results <- test_data %>% 
  select(price_category) %>% 
  bind_cols(pdx_fit %>% 
              predict(new_data = test_data)) %>% 
  bind_cols(pdx_fit %>% 
              predict(new_data = test_data, type = "prob"))

# Print the results
pred_results %>% 
  slice_head(n = 10)
```

Let's take a look at the confusion matrix:

```{r}
pred_results%>% 
  conf_mat(price_category, .pred_class) %>% 
  autoplot(type = "heatmap")+
  theme_minimal()
```

# Evaluate other desired metrics

```{r}

eval_metrics <- metric_set(ppv, recall, accuracy, f_meas)
eval_metrics(data = pred_results, truth = price_category, estimate = .pred_class)
```

```{r}
pred_results %>% 
  roc_auc(price_category, .pred_above)
```

### Make a roc_chart

Let's create the ROC curve. Again, since the event we are predicting is the first level in the price_category factor ("above"), we provide roc_curve() with the relevant class probability .pred_above:

```{r}
pred_results %>% 
  roc_curve(truth = price_category, .pred_above) %>% 
  autoplot()+
  theme_minimal()
```

## Reference

Tidymodels- https://www.tidymodels.org/
